---
output: github_document
bibliography: docs/arulesCBA.bib
link-citations: yes
---

```{r echo=FALSE, results = 'asis'}
pkg <- 'arulesCBA'

source("https://raw.githubusercontent.com/mhahsler/pkg_helpers/main/pkg_helpers.R")
pkg_title(pkg)
```

The R package [arulesCBA](https://cran.r-project.org/package=arulesCBA) [@Hahsler.Johnson.ea:2019]
is an extension of the package [arules](https://cran.r-project.org/package=arules) [@Hahsler.Grün.ea:2005] to perform
association rule-based classification. The package provides the infrastructure for class association rules and implements associative classifiers based on the following algorithms:

* __CBA__:    Classification Based on Association Rules [@Liu.Hsu.ea:1998].
* __CMAR__:   Classification based on Multiple Association Rule  [@Li.Han.ea:2001] via LUCS-KDD Software Library.
* __CPAR__:   Classification based on Predictive Association Rules [@Yin.Han:2003] via LUCS-KDD Software Library.
* __C4.5__:   Rules extracted from a C4.5 decision tree [@Quinlan:1993] via J48 in R/Weka.
* __FOIL__:   First-Order Inductive Learner [@Quinlan.Cameron-Jones:1993;@Yin.Han:2003].
* __PART__:   Rules from Partial Decision Trees [@Frank.Witten:1998] via R/Weka.
* __PRM__:    Predictive Rule Mining [@Yin.Han:2003] via LUCS-KDD Software Library.
* __RCAR__:   Regularized Class Association Rules using Logistic Regression [@Azmi.Runger.ea:2019].
* __RIPPER__: Repeated Incremental Pruning to Produce Error Reduction [@Cohen:1995] via R/Weka.

The package also provides the infrastructure for associative classification (supervised discetization, mining class association rules (CARs)), and implements various association rule-based classification strategies
(first match, majority voting, weighted voting, etc.).
Some algorithms are interfaced by the R package R/Weka [@Hornik.Buchta.ea:2009] and the LUCS-KDD Software Library [@Coenen:2013].


```{r echo=FALSE, results = 'asis'}
pkg_usage(pkg)
```

```{r echo=FALSE, results = 'asis'}
pkg_citation(pkg, 2)
```       


```{r echo=FALSE, results = 'asis'}
pkg_install(pkg)
```
## Usage

```{r}
library("arulesCBA")
data("iris")
```

Learn a classifier.

```{r}
classifier <- CBA(Species ~ ., data = iris)
classifier
```

Inspect the rulebase.

```{r}
inspect(classifier$rules, linebreak = TRUE)
```
  
Make predictions for the first few instances of iris.

```{r}
predict(classifier, head(iris))
```

## References

